---
title: "LI water temps 2019"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(plotly)
library(tidyverse)
library(dplyr)
library(magrittr)

library(chron)
library(stringr)
library(htmltools)

library(lubridate)
library(dataRetrieval)

# plotting tools
library(ggplot2)
library(ggthemes)
library(ggExtra)
library(scales)
library(hrbrthemes)
# Spatial packages

library(sf)
library(maptools)
library(RColorBrewer)
library(mapview)
library(knitr)
library(rmarkdown)

library(weatherData)
library(lunar)

# timeseries 
library(tsibble)
library(fable)
library(feasts)
library(xts)

```

### Water temperature stations

Exploring temperature data in the Peconics.
2 USGS guage stations exist with continuous data collection. 


```{r sites}
h5("Data presented within this document have been made available through the",
      a(href = 'http://waterdata.usgs.gov/nwis', "USGS NWIS Web Interface:")
    )
# estuarySites <- whatNWISsites(stateCd = "NY", 
#               # water temp C
#               parameterCd = "00010") %>% 
#   filter(site_tp_cd == "ES") %>% 
#   pull(site_no)

peconicSiteIDs <- c("01304200", "01304562")


dailyTemps <- readNWISdv(siteNumbers = peconicSiteIDs, 
                         startDate = "2013-01-01", #first full year of data collection 2013.
                         endDate = "2019-11-01",
                         statCd = c("00001", "00002", "00003"),
                         parameterCd = "00010") %>% 
  dataRetrieval::renameNWISColumns() 
dailyDOs <- readNWISdv(siteNumbers = peconicSiteIDs, 
                         startDate = "2013-01-01", #first full year of data collection 2013.
                         endDate = "2019-11-01",
                         statCd = c("00001", "00002", "00003"),
                         parameterCd = c("00300", "00301")) %>% 
  dataRetrieval::renameNWISColumns() 

dailyTemps <- dailyTemps %>% filter(Wtemp < 100 & Wtemp > -10) # filter out coded -9999s and 9999s


estuarySiteLocations <- readNWISsite(peconicSiteIDs) %>% 
  st_as_sf(coords = c("dec_long_va", "dec_lat_va"), crs = 4269) %>% 
  mutate(stationName = case_when(
    station_nm == "ORIENT HARBOR AT ORIENT NY" ~ "Orient Harbor",
    station_nm == "PECONIC RIVER AT COUNTY HWY 105 AT RIVERHEAD NY" ~ "Peconic River Rte 105"
  ))

# popup 

	popupLink <- function(siteName, siteNo){
		url <- paste('<a href="https://waterdata.usgs.gov/nwis/inventory?agency_code=USGS&site_no=', 
			     siteNo,'"','> Access data and site information from NWIS </a>', sep = "")
		
		popupFinal <- paste(sep = "<br/>", 
				    siteName, 
				    url)
		popupFinal
	}

stationMap <- mapview(estuarySiteLocations, popup = popupLink(siteName = estuarySiteLocations$stationName, siteNo = estuarySiteLocations$site_no))
cntrcords <- c(mean(st_coordinates(estuarySiteLocations)[,1]),
               mean(st_coordinates(estuarySiteLocations)[,2]))
stationMap@map %>% leaflet::setView(cntrcords[1], cntrcords[2], zoom = 10)

```

## Plots
 
Data has been collected since 2012, with 2013 being the first 'full' year of data collection.


```{r temps, echo=FALSE}
# go tsibble.
theme_set(theme_ipsum())

temps <- dailyTemps %>% 
  # join to get station names
  left_join(estuarySiteLocations %>% 
              select(site_no, stationName)) %>%
   as_tsibble(key = stationName, index = Date) %>% 
  fill_gaps() %>% 
  mutate(month = month(Date),
         year = year(Date),
         day_of_year = yday(Date),
         commonDate = as.Date(paste0("2000-",format(Date, "%j")), "%Y-%j"),
         year2019 = ifelse(year == 2019, 1, 0))


temps  %>% 
  ggplot(aes(x = commonDate, y = Wtemp, color = year2019, group = year)) +
  geom_line() + 
  scale_color_gradient(low = "grey", high = "blue", aesthetics = "color") +
  # scale_color_viridis_c(option = "inferno") +
  # facet_grid(stationName ~ year) +
  facet_grid(stationName ~ .) +
  scale_x_date(labels = function(x) format(x, "%d-%b")) +
  theme(panel.spacing = unit(.5, "lines")) +
  # scale_x_continuous(labels = function(x) format(as.Date(as.character(x), "%j"), "%d-%b")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(y = "Temp (c)", x = NULL, title = "Water temperature through the year", subtitle = "2019 shown in blue")


```

## Anomolies?  
How do the temps of 2019 compare with the past? 
Red ribbon = 2019 temperature range based on daily max and min. Grey ribbon is *daily average* max and min by day for the previous 6 years. 
 
```{r anomolies, message=FALSE, warning=FALSE, paged.print=FALSE}
daily_mean_6yr <- temps %>% 
  # filter to get 2012-2018
  filter(year != 2019) %>% 
  # group by site (site is key)
  group_by_key() %>%
  # index by the day of year
  index_by(day_of_year = ~ yday(.)) %>% 
  summarize(temp_mean7yr = mean(Wtemp, na.rm = TRUE),
            temp_max7yr = mean(Wtemp_Max, na.rm = TRUE),
            temp_min7yr = mean(Wtemp_Min, na.rm = TRUE)) %>% 
  mutate(commonDate = as.Date(day_of_year, origin = "2019-01-01"))

anomPlot <- daily_mean_6yr %>% ggplot(aes(x = commonDate, y = temp_mean7yr)) + 
  geom_ribbon(aes(ymax = temp_max7yr, ymin = temp_min7yr), alpha = 0.6) +
  # geom_line(aes(y = temp_min7yr)) +
  # geom_line(aes(y = temp_max7yr)) + 
  geom_ribbon(data = temps %>% filter(year == 2019), 
              aes(y = Wtemp, ymin = Wtemp_Min, ymax = Wtemp_Max, x = Date), fill = 'red', alpha = 0.6) + 
  # theme_minimal() + 
  scale_x_date(date_labels = "%d-%b", date_breaks = "1 month", limits = c(ymd("2019-01-01"), Sys.Date())) +
  # scale_x_date(labels = function(x) format(x, "%d-%b")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_grid(stationName ~ .) + theme(panel.spacing = unit(1, "lines")) +
  labs(main = "2019 Daily temperatures compared to 6 year means (2013-2018)", 
       x = NULL, y = "Degrees C")

anomPlot

```

### Summer temps
Same data as above, with a focus on the summer season.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
anomPlot + 
  scale_x_date(date_labels = "%d-%b", date_breaks = "2 week", limits = c(ymd("2019-06-01"), ymd("2019-09-01")))
```

### Dissolved Oxygen 
Similar assessment as above, focus on dissolved oxygen

```{r}

oxygen <- dailyDOs %>% 
  # join to get station names
  left_join(estuarySiteLocations %>% 
              select(site_no, stationName))  %>%  
  select(-geometry) %>% 
   as_tsibble(key = stationName, index = Date) %>% 
  fill_gaps() %>% 
  mutate(month = month(Date),
         year = year(Date),
         day_of_year = yday(Date),
         commonDate = as.Date(paste0("2000-",format(Date, "%j")), "%Y-%j"),
         year2019 = ifelse(year == 2019, 1, 0))

 


oxygen  %>% 
  ggplot(aes(x = commonDate, y = DO, color = year2019, group = year)) +
  geom_line() + 
  scale_color_gradient(low = "grey", high = "red", aesthetics = "color") +
  # scale_color_viridis_c(option = "inferno") +
  # facet_grid(stationName ~ year) + 
  facet_grid(stationName ~ .) + 
  scale_x_date(labels = function(x) format(x, "%d-%b")) +
  theme(panel.spacing = unit(.5, "lines")) +
  # scale_x_continuous(labels = function(x) format(as.Date(as.character(x), "%j"), "%d-%b")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(y = "DO (mg/L)", x = NULL)



```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
daily_mean_oxygen_6yr <- oxygen %>% 
  # filter to get 2012-2018
  filter(year != 2019) %>% 
  # group by site (site is key)
  group_by_key() %>%
  # index by the day of year
  index_by(day_of_year = ~ yday(.)) %>% 
  summarize(oxygen_mean6yr = mean(DO, na.rm = TRUE),
            oxygen_max6yr = mean(DO_Max, na.rm = TRUE),
            oxygen_min6yr = mean(DO_Min, na.rm = TRUE)) %>% 
  mutate(commonDate = as.Date(day_of_year, origin = "2019-01-01"))

daily_mean_oxygen_6yr %>% ggplot(aes(x = commonDate, y = oxygen_mean6yr)) + 
  geom_ribbon(aes(ymax = oxygen_max6yr, ymin = oxygen_min6yr), alpha = 0.6) +
  geom_ribbon(data = oxygen %>% filter(year == 2019), 
              aes(y = DO, ymin = DO_Min, ymax = DO_Max, x = Date), fill = 'red', alpha = 0.6) + 
  # theme_minimal() + 
  scale_x_date(date_labels = "%d-%b") +
  # scale_x_date(labels = function(x) format(x, "%d-%b")) +
  facet_grid(stationName ~ .) + theme(panel.spacing = unit(1, "lines")) +
  labs(main = "2019 Daily Mean Dissolved Oxygen measures compared to 6 year means (2013-2018)", 
       x = NULL, y = "DO (mg/L)")


```






