---
title: "Data sketches of water temperatures and thresholds"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})

output: html_document
---

```{r setup, include=FALSE}
here::i_am("waterTempThresholds.Rmd")

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(plotly)
library(tidyverse)
library(dplyr)
library(magrittr)
library(glue)

library(chron)
library(stringr)
library(htmltools)

library(lubridate)
library(dataRetrieval)

# plotting tools
library(ggplot2)
library(ggthemes)
library(ggExtra)
library(ggtext)
library(gghighlight)
library(scales)
library(hrbrthemes)
# Spatial packages

library(sf)
library(maptools)
library(RColorBrewer)
library(mapview)
library(knitr)
library(rmarkdown)

# library(weatherData)
library(lunar)
# Tidal data
library(VulnToolkit) 
library(Tides)

# timeseries 
library(tsibble)
library(fable)
library(feasts)
library(xts)
library(here)
library(units)

source(here::here("../LI_SaltMarshMonitoring/R/truBlue.R")) # lots of code went into trublue utilities from another project. consider making a package.
library(lunar)
library(suncalc)
library(extrafont)

```

### Water temperature stations

Exploring temperature data in the Peconics.
2 USGS guage stations exist with continuous data collection. 


```{r sites}
h5("Data presented within this document have been made available through the",
      a(href = 'http://waterdata.usgs.gov/nwis', "USGS NWIS Web Interface:")
    )

tempsSites <- whatNWISsites(stateCd = "NY",
              # water temp C
              parameterCd = "00010", 
              hasDataTypeCd = "dv", 
              siteStatus = "active") %>%
  filter(site_tp_cd == "ES") %>%
  pull(site_no)




dailymeanTemps <- readNWISdv(
  siteNumbers = tempsSites,
  statCd = "00003",
  startDate = "2013-01-01",
  #first full year of data collection 2013.
  endDate = Sys.Date(),
  # statCd = c("00001", "00002", "00003", "00009"), #
  parameterCd = "00010"
) %>%
  dataRetrieval::renameNWISColumns() %>% 
  select(agency_cd:Wtemp_cd)


dailymeanTemps <- dailymeanTemps %>% filter(Wtemp < 100 & Wtemp > -10) # filter out coded -9999s and 9999s


tempsSitesLocations <- readNWISsite(tempsSites) %>% 
  st_as_sf(coords = c("dec_long_va", "dec_lat_va"), crs = 4269) %>% 
  mutate(stationName = case_when(
    station_nm == "ORIENT HARBOR AT ORIENT NY" ~ "Orient Harbor",
    station_nm == "PECONIC RIVER AT COUNTY HWY 105 AT RIVERHEAD NY" ~ "Peconic River Rte 105",
    station_nm == "FROST CREEK AT SHEEP LN BRIDGE AT LATTINGTOWN NY" ~ "Frost Creek",
    station_nm == "EAST CREEK AT SANDS POINT NY" ~ "East Creek",
    station_nm == "GREAT SOUTH BAY AT WEST SAYVILLE NY" ~ "West Sayville",
    station_nm == "REYNOLDS CHANNEL AT POINT LOOKOUT NY" ~ "Point Lookout",
    station_nm == "HOG ISLAND CHANNEL AT ISLAND PARK NY" ~ "Hog Island",
    station_nm == "ROCKAWAY INLET NEAR FLOYD BENNETT FIELD NY" ~ "Rockaway Inlet", 
    station_nm == "HUDSON RIVER BELOW POUGHKEEPSIE NY" ~ "Poughkeepsie", 
    station_nm == "HUDSON RIVER AT SOUTH DOCK AT WEST POINT NY" ~ "West Point", 
    station_nm == "HUDSON RIVER AT PIERMONT NY" ~ "Piermont",
    station_nm == "FLAX POND AT OLD FIELD NY" ~ "Flax Pond",
    station_nm == "SHELTER ISLAND SOUND AT SHELTER ISLAND NY" ~ "Shelter Is Sound",
    station_nm == "HUDSON RIVER NEAR POUGHKEEPSIE NY" ~ "Hudson Rv Poughkeepsie",
    T ~ station_nm
    
  ))

# popup 

	popupLink <- function(siteName, siteNo){
		url <- paste('<a href="https://waterdata.usgs.gov/nwis/inventory?agency_code=USGS&site_no=', 
			     siteNo,'"','> Access data and site information from NWIS </a>', sep = "")
		
		popupFinal <- paste(sep = "<br/>", 
				    siteName, 
				    url)
		popupFinal
	}

stationMap <- mapview(tempsSitesLocations, popup = popupLink(siteName = tempsSitesLocations$stationName, siteNo = tempsSitesLocations$site_no))
cntrcords <- c(mean(st_coordinates(tempsSitesLocations)[,1]),
               mean(st_coordinates(tempsSitesLocations)[,2]))
stationMap@map %>% leaflet::setView(cntrcords[1], cntrcords[2], zoom = 8)



```

## Plots
 
Data has been collected since 2012, with 2013 being the first 'full' year of data collection.


```{r tempdata, echo=FALSE}
# go tsibble.
theme_set(theme_ipsum())

dailymeanTemps_clean <- dailymeanTemps %>% 
  # join to get station names
  left_join({tempsSitesLocations %>% 
              st_drop_geometry() %>% 
              select(site_no, stationName)}, by = "site_no") %>%
  as_tsibble(key = site_no, index = Date) %>% 
  fill_gaps() %>%
  mutate(month = month(Date),
         year = year(Date),
         day_of_year = yday(Date),
         commonDate = as.Date(paste0("2000-",format(Date, "%j")), "%Y-%j")) %>% 
  # Calc ecdf first -
  group_by(stationName, year)

```


```{r tempplot, echo=FALSE}

dailymeanTemps_clean  %>% 
  ggplot(aes(x = commonDate, y = Wtemp, group = year)) +
  geom_line() + 
  # gghighlight(year == 2013) +
  # scale_color_gradient(low = "grey", high = "blue", aesthetics = "color") +
  # scale_color_viridis_c(option = "inferno") +
  # facet_grid(stationName ~ year) +
  facet_grid(stationName ~ .) +
  scale_x_date(labels = function(x) format(x, "%d-%b")) +
  theme(panel.spacing = unit(.5, "lines")) +
  # scale_x_continuous(labels = function(x) format(as.Date(as.character(x), "%j"), "%d-%b")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1),
        strip.text.y = element_text(angle = 0)) #+ 
  # labs(y = "Temp (c)", x = NULL, title = "Water temperature through the year", subtitle = "2019 shown in blue")


```


```{r}


dailymeanTemps_clean  %>% 
  ggplot(aes(x = commonDate, y = Wtemp, group = year)) +
  geom_line(aes(color = Wtemp > 25)) + 
  # gghighlight(year) +
  # scale_color_gradient(low = "grey", high = "blue", aesthetics = "color") +
  scale_color_viridis_d(option = "inferno") +
  # facet_grid(stationName ~ year) +
  facet_grid(stationName ~ .) +
  scale_x_date(labels = function(x) format(x, "%d-%b")) +
  theme(panel.spacing = unit(.5, "lines")) +
  # scale_x_continuous(labels = function(x) format(as.Date(as.character(x), "%j"), "%d-%b")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1),
        strip.text.y = element_text(angle = 0)) + 
  labs(y = "Temp (c)", x = NULL, title = "Water temperature through the year", subtitle = "highlights above 25C")



```




### Exploring Degree Hours - or Degree Days - 
Although the upper limit of zostera is often the focus, the growing degree days typically captures the temps above a reference baseline - which is typically the lower physiological limit of the plant (the temperature it starts growing at).

### Start with Growing Degree Days.

Chillr has a check data function that looks at the data object for compliance with the funcitons. It seems that they use the Tmax - Tmin /2. That could  be just to allow the ability to use that approach instead of the Tmean as some methods use. More to explore.


```{r}
temp_threshold <- 28
```


```{r}


dailymeanTemps_clean %>%
  filter(stationName %in% c("West Sayville")) %>%
  filter(year != 2023) %>% 
  as_tibble() %>%
  group_by(year, stationName) %>%
  arrange(stationName, Date) %>% 
  mutate(
    threshold = temp_threshold,
    threshold_over = Wtemp > threshold,
    # returns boolean 1 for YES it is greater than threshold
    deg_day = (Wtemp - threshold) * threshold_over,
    accumulated_gdd = cumsum(deg_day)
  ) %>% 
  group_by(stationName, commonDate) %>% 
  mutate(mean_gdd = mean(accumulated_gdd)) %>% 
  group_by(year, stationName) %>%
  rowwise() %>% 
  mutate(anomoly_gdd = accumulated_gdd - mean_gdd) %>% 

  ggplot(aes(x = commonDate, y = deg_day, group = year, color = as.factor(year))) +
  geom_line(size = 0.85) +
  facet_grid(year ~ stationName ) +
  scale_x_date(
    labels = function(x)
      format(x, "%d-%b")
  ) +
  theme(panel.spacing = unit(.5, "lines")) +
  # scale_x_continuous(labels = function(x) format(as.Date(as.character(x), "%j"), "%d-%b")) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.text.y = element_text(angle = 0)
  ) +
  labs(
    y = "GDD degC",
    color = "",
    x = NULL,
    title = glue("Growing Degree Days"),
    subtitle = glue("referenced to base line of {temp_threshold}")
  )

ggplotly()


```


## Anomoly Look
```{r}

dailymeanTemps_clean %>%
  filter(stationName %in% c("West Sayville")) %>%
  filter(year != 2023) %>% 
  as_tibble() %>%
  group_by(year, stationName) %>%
  arrange(stationName, Date) %>% 
  mutate(
    threshold = temp_threshold,
    threshold_over = Wtemp > threshold,
    # returns boolean 1 for YES it is greater than threshold
    deg_day = (Wtemp - threshold) * threshold_over,
    accumulated_gdd = cumsum(deg_day)
  ) %>% 
  group_by(stationName, commonDate) %>% 
  mutate(mean_gdd = mean(accumulated_gdd)) %>% 
  group_by(year, stationName) %>%
  rowwise() %>% 
  mutate(anomoly_gdd = accumulated_gdd - mean_gdd) %>% 

  ggplot(aes(x = commonDate, y = anomoly_gdd, group = year, color = as.factor(year))) +
  geom_line(size = 0.85) +
  facet_grid(stationName ~ .) +
  scale_x_date(date_breaks = "1 week",
    labels = function(x)
      format(x, "%d-%b")
  ) +
  theme(panel.spacing = unit(.5, "lines")) +
  # scale_x_continuous(labels = function(x) format(as.Date(as.character(x), "%j"), "%d-%b")) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.text.y = element_text(angle = 0)
  ) +
  labs(
    y = "GDD degC",
    color = "",
    x = NULL,
    title = glue("Growing Degree Days - anomoly from mean GDD of temperature record"),
    subtitle = glue("referenced to base line of {temp_threshold}")
  )

ggplotly()

```






